---
title: "Project title"
subtitle: "Exploratory data analysis"
format: html
editor: visual
execute:
  echo: true
---

# Research question(s)

Research question(s). State your research question (s) clearly.

RQ: What are the countries that experienced the most drastic shift in educational attainability among the females and the lower classes. What countries did not experience that much of a shift?

# Data collection and cleaning

Have an initial draft of your data cleaning appendix. Document every step that takes your raw data file(s) and turns it into the analysis-ready data set that you would submit with your final project. Include text narrative describing your data collection (downloading, scraping, surveys, etc) and any additional data curation/cleaning (merging data frames, filtering, transformations of variables, etc). Include code for data curation/cleaning, but not collection.

```{r}
#|: ex1

library(tidyverse)
library(skimr)
library(janitor)

soccer <- read_csv("data/top_goalscorers.csv")
soccer <- clean_names(soccer)

soccer1 <- soccer |>
  select(-c(games_number, goals_saves, dribbles_past, penalty_won, penalty_commited))
```

John: explain how you got the data

# Data description

Have an initial draft of your data description section. Your data description should be about your analysis-ready data.

# Data limitations

Identify any potential problems with your dataset.

# Exploratory data analysis

Perform an (initial) exploratory data analysis.

```{r}
#|: exploratory-data-analysis 

library(tidyverse)
library(tidymodels)

#Example From ae14
# Let's start by taking a look at our data. Create an density plot to investigate the relationship between `spam` and `exclaim_mess`. Additionally, calculate the mean number of exclamation points for both spam and non-spam emails.

# ggplot(data = email, mapping = aes(x = exclaim_mess, fill = spam)) +
#   geom_density(alpha = 0.5) + 
#   scale_x_sqrt()
```

# Questions for reviewers

List specific questions for your peer reviewers and project mentor to answer in giving you feedback on this phase.
