---
title: "Project title"
subtitle: "Exploratory data analysis"
format: html
editor: visual
execute:
  echo: true
---

# Research question(s)

Research question(s). State your research question (s) clearly.

Is there a correlation between the height and/or weight and specific game statistics (Including: goals scored, assists, penalties, passes, etc.)

-   Do teams with taller players tend to have more penalties?

Are some leagues better than others?

-   Are some leagues outputting more goals than others?

-   Which team with the greatest goal output in every league has the most penalties (i.e. yellow/red cards)?

# Data collection and cleaning

Have an initial draft of your data cleaning appendix. Document every step that takes your raw data file(s) and turns it into the analysis-ready data set that you would submit with your final project. Include text narrative describing your data collection (downloading, scraping, surveys, etc) and any additional data curation/cleaning (merging data frames, filtering, transformations of variables, etc). Include code for data curation/cleaning, but not collection.

```{r}
#|: ex1

library(tidyverse)
library(skimr)
library(janitor)

soccer <- read_csv("data/top_goalscorers.csv")
soccer <- clean_names(soccer)

soccer1 <- soccer |>
  select(-c(games_number, goals_saves, dribbles_past, penalty_won, penalty_commited))
```

John: explain how you got the data

# Data description

Have an initial draft of your data description section. Your data description should be about your analysis-ready data.

```{r}
View(soccer1)
glimpse(soccer1)
```

-   What are the observations (rows) and the attributes (columns)?

    -   There are 560 observations and 56 attributes (columns). The dataset contains player statistics (ex/ number of appearances, number of minutes played, number of yellow or red cards attained) for the top 20 scorers from the top 27 football leagues for the 2021 league season.

-   Why was this dataset created?

    -   This dataset was created to rank and compare the top scoring football players across top leagues. This could be helpful for football fans, sports betters, or data scientists!

-   Who funded the creation of the dataset?

    -   N/a (Most stats are automatically recorded, but were probably scraped from some summary site)

-   What processes might have influenced what data was observed and recorded and what was not?

    -   It may have been influenced by referee rulings, as some scores may not have counted in the final game statistics.

-   What preprocessing was done, and how did the data come to be in the form that you are using?

    -   ***JOHN:***

-   If people are involved, were they aware of the data collection and if so, what purpose did they expect the data to be used for?

    -   The people involved are football players, who are aware of data collection, a this occurs during and after all of their games.

    -   They probably expected the data to be used for:

        -   general comparison (past/future versions of themselves, comparing their team to another, or between leagues)

        -   sports reporting (to football fans)

        -   MVP decisions

# Data limitations

-   Data is only from the 2021 season. Therefore, we cannot make comparisons within groups (player, team, or league) over time.

-   There are no goalies in the dataset (best scorers i.e. offensive players) therefore, we are unable to assess a team's defense, which is important when considering how good a team is!

# Exploratory data analysis

Perform an (initial) exploratory data analysis.

```{r}
#|: exploratory-data-analysis 

library(tidyverse)
library(tidymodels)

#Example From ae14
# Let's start by taking a look at our data. Create an density plot to investigate the relationship between `spam` and `exclaim_mess`. Additionally, calculate the mean number of exclamation points for both spam and non-spam emails.

# ggplot(data = email, mapping = aes(x = exclaim_mess, fill = spam)) +
#   geom_density(alpha = 0.5) + 
#   scale_x_sqrt()
```

# Questions for reviewers

-   How many research questions should we have-- if only one or two, which ones do you suggest?

-   The data we are using is already pretty tidy (tidied when converting from API --\> JSON --\> CSV), therefore there are minimal cleaning steps in R\-- confirming that this is acceptable, as it was indicated in the proposal directions that using an API was OK.
