---
title: Soccer Stats Across Leagues
author: Skillful Pikachu <br> Henry Friedberg, Pranathi Adhikari, John Aclufi, Shara Shanmugasundaram 
title-slide-attributes:
  style: |
    .title {
      font-size: 50px;
    }
    .author {
      font-size: 40px;
    }
date: May 5, 2023
format: 
  revealjs:
    theme: "serif"
editor: visual
execute:
  echo: false
  cache: true
---

```{r}
#| label: setup
#| include: false
#| cache: false
library(tidyverse)
library(rcis)
library(scales) # formatting numbers

theme_set(theme_minimal(base_size = 16))
```

## Topic and Motivation

<br>

**Topic:** Analyze offensive stats for soccer players in various leagues during 2021 season.

**Motivation:** Find relationships between soccer players, leagues, and their stats.

<br>

**Research Questions:**

-   What factors influence booking frequency of players?
-   Does aggressiveness differ between European and non-European leagues?

## Data Introduction {.smaller}

::: columns
::: {.column width="50%"}
```{r}
#| label: data-introduction
#| message: false
#| warning: false
library(tidyverse)
library(skimr)
library(dplyr)
library(kableExtra)

soccer <- read_csv("data/top_goalscorers.csv")

soccer_display <-soccer |>
  select("player.name", "player.age", "goals.total") |>
  slice(1:5) |>
  rename("Player Name" = player.name, "Player Age" = player.age,
         "Goals Scored" = goals.total)


kable(soccer_display)

```
:::

::: {.column width="50%"}
<div>

-   Dataset contains the top 20 scorers from the top 27 football leagues for the 2021 league season.

    -   Rows: Individual Players

    -   Columns: Attributes

-   The data is used by football fans, sports bettors, and researchers

</div>
:::
:::

## Highlights from EDA

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Inference/modeling/other analysis

Your results here

## Inference/modeling/other analysis

Your results here

## Conclusions

#### RQ1 Conclusion

-   What factors influence booking frequency of players?

    -   **We determined that rating was the only significant predictor of bookings in our analysis**

#### RQ2 Conclusion

-   Does aggressiveness differ between European and non-European leagues?

    -   **We determined that only average goals per minute and total dues are significant predictors when determining the league.**
